---
title: "In-class Exercise 4"
author: "Dr. Kam Tin Seong"
date: "5 May 2023"
date-modified: "`r Sys.Date()`"
execute:
  echo: true
  eval: true
  warning: false
---

```{r}
pacman::p_load(rstatix, gt, patchwork, tidyverse)
```

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

Visualising Normal Distribution: 

A Q-Q plot, short for “quantile-quantile” plot, is used to assess whether or not a set of data potentially came from some theoretical distribution.  In most cases, this type of plot is used to determine whether or not a set of data follows a normal distribution.

If the data is normally distributed, the points in a Q-Q plot will lie on a straight diagonal line. Conversely, if the points deviate significantly from the straight diagonal line, then it’s less likely that the data is normally distributed.

::: panel-tabset
## The plot
```{r}
#| echo: false
#| fig-width: 4
#| fig-height: 4
ggplot(exam_data, 
       aes(sample=ENGLISH)) +
  stat_qq() + 
  stat_qq_line()
```

## The code chunk
```{r}
#| eval: false
#| fig-width: 4
#| fig-height: 4
ggplot(exam_data, 
       aes(sample=ENGLISH)) +
  stat_qq() + 
  stat_qq_line()
```
:::

::: {.callout-note}
We can see that the points deviate significantly from the straight diagonal line. This is a clear indication that the set of data is not normally distributed.
:::

### Combining statistical graph and analysis table

Need to install webshot

::: panel-tabset

## The plot
```{r}
#| echo: false
qq <- ggplot(exam_data, 
       aes(sample=ENGLISH)) +
  stat_qq() + 
  stat_qq_line()

sw_t <- exam_data %>%
  shapiro_test(ENGLISH) %>%
  gt()

tmp <- tempfile(fileext = '.png') 
gtsave(sw_t, tmp)
table_png <- png::readPNG(tmp, native = TRUE)

qq + table_png
```

## The code chunk
```{r}
#| eval: false
qq <- ggplot(exam_data, 
       aes(sample=ENGLISH)) +
  stat_qq() + 
  stat_qq_line()

sw_t <- exam_data %>%
  shapiro_test(ENGLISH) %>%
  gt()

tmp <- tempfile(fileext = '.png') 
gtsave(sw_t, tmp)
table_png <- png::readPNG(tmp, native = TRUE)

qq + table_png
```
:::

### Facet version

::: panel-tabset

## The plot
```{r}
#| echo: false
#| fig-width: 12
#| fig-height: 12
ggplot(exam_data, 
       aes(sample=ENGLISH)) +
  stat_qq() + 
  stat_qq_line() +
  facet_wrap(~ CLASS,
             nrow = 3,
             ncol = 3)
```

## The code chunk
```{r}
#| eval: false
ggplot(exam_data, 
       aes(sample=ENGLISH)) +
  stat_qq() + 
  stat_qq_line() +
  facet_wrap(~ CLASS,
             nrow = 3,
             ncol = 3)
```
:::



